Psychophysics of musical elements in the discrete-time representation of sound

### 2013-09-19 article/article.tex
feito ateh I. A

Music is commonly defined as the art made by sounds and silences. Sound is the longitudinal wave of mechanical pressure. The frequency bandwidth between $20Hz$ and $20kHz$ is appreciated by the human hearing system with boundaries dependent on the person, climate conditions and sonic characteristics themselves~\cite{Roederer}. Considered the speed of sound as $\approx 343.2 m/s$, these limits corresponds to wavelengths of $\frac{343.2}{20} = 17.16\,m$ and $\frac{343.2}{20000}=17.16\,mm$.

Human perception of sound involves captivation by bones, stomach, ears, transfer functions of head and processing torso and nervous system~\cite{}. Besides that, the ear is a dedicated organ responsible for the capture of these waves. Its mechanism decomposes sound into its sinusoidal spectrum and delivers them to the nervous system. The sinusoidal components are crucial to musical phenomena, as one can perceive the composition of sounds according to a specific musical interest, e.g., tunings and scales.

The representation of sound is commonly called audio (although these terms are often used without distinction). Audio expresses the waveform prominent from the capture of microphones or from synthesis. In order to account for the corresponding digital representation, digital audio is specified by protocols that eases file storage and transferring, usually with low loss of quality. Standard representation of digital audio consists of equally spacing samples by $\lambda_s$ durations in time, where each sample is specified by a sample number of bits. This process is called the Pulse Code Modulation (PCM) representation of sound. A PCM digital sound is characterized by it's sampling frequency $f_s=\frac{1}{\lambda_s}$ (also called the sampling rate), and by a bit depth, which is the number of bits used for representing the amplitude of each sample. With the purposes of illustrating this idea, Figure~\ref{fig:PCM} shows $25$ samples of a PCM audio with $4$ bits each. The $2^4=16$ possible steps for each sample, together with the regular spacing $\lambda_s$ between them, establish what is named as quantization error. The noise prominent of these errors diminishes conforming the spacing also does.

@@

The Nyquist theorem~\cite{} says that half of the sampling frequency is the maximum frequency of the signal. Thus, it is necessary to have a sampling frequency at least twice the highest frequency heard by humans, that is, $f_s \geq 2\times 20kHz = 40kHz$. This configures the basis for the use of sampling frequencies like $44.1kHz$ and $48kHz$, standards in Compact Disks (CD) and Broadcast systems (Radio and TV), respectively.

For the average listener -- and a reasonable part of specialists -- the notion of music presupposes rhythmic and pitch organization. I a simple  way, rhythm can be described as recurring pattern produced by notes differing in duration, pause (silence) and intensity. Pitch configures the perceptual attribute that allows sounds to be organized in a frequency-related scale~\cite{}. Music from the twentieth century enlarged this traditional comprehension of music. This occurred in the concert music, specially in the concrete, and in both electronic and electroacoustic styles. On the last decade of the century, it became evident that popular music had also incorporated sounds without defined pitch and temporal organization, to name a few of simple metrics. Even though, the note still stands paradigmatic as a 'fundamental unit' of musical structures and, in practice, it can unfold in sounds that observe this recent developments (**not clear here**).

@@

\nocite{*}
\bibliography{phi+mus}
\end{document}

>>>>

Music is commonly defined as the art made by sounds and silences. Sound is the longitudinal wave of mechanical pressure. The frequency bandwidth between $20Hz$ and $20kHz$ is appreciated by the human hearing system with boundaries dependent on the person, climate conditions and sonic characteristics themselves~\cite{Roederer}. Considered the speed of sound as $\approx 343.2 m/s$, these limits imply wavelengths of $\frac{343.2}{20} = 17.16\,m$ and $\frac{343.2}{20000}=17.16\,mm$.

Human perception of sound involves captivation by bones, stomach, ears, transfer functions of head and torso, and processing by the nervous system~\cite{}. Besides that, the ear is a dedicated organ for the appreciation of these waves. Its mechanism decomposes sound into its sinusoidal spectrum and delivers to the nervous system. The sinusoidal components are crucial to musical phenomena, as one can perceive in the composition of sounds of musical interest (such as harmonic sounds and noises, seen in sections~\ref{sec:discNote} and~\ref{sec:internalVar}), and sonic structures of musical interest (such as tunings and scales, in section~\ref{sec:notesMusic}).

The representation of sound is commonly called audio (although these terms are often used without distinction). Audio expresses waves prominent from the capture of microphones or from direct synthesis, although these sources entangle as captured sounds are processed to generate new sonorities. Digital audio specified by protocols that eases file transfering and storage often imply a quality loss. Standard representation of digital audio, on the other hand, assures perfect reconstruction of the analog wave, within any convenient precision. This paradigm consists of equally spaced samples, by $\lambda_s$ durations, each specified by a fixed number of bits. This is the Pulse Code Modulation (PCM) representation of sound. A sound in PCM audio is characterized by a sampling frequency $f_s=\frac{1}{\lambda_s}$ (also called the sampling rate), which is the number of samples used for representing a second of sound; and by a bit depth, which is the number of bits used for representing the amplitude of each sample. Figure~\ref{fig:PCM} shows $25$ samples of a PCM audio with $4$ bits each. The fixed $2^4=16$ values for the amplitude of each sample, with the regular spacing $\lambda_s$, yields the quantization error or noise. This noise diminishes as bit depth increases.

@@

The Nyquist theorem~\cite{} states that the sampling frequency is twice the maximum frequency of the represented signal. Thus, for general musical purposes, it is necessary to have samples in a rate at least twice the highest frequency heard by humans, that is, $f_s \geq 2\times 20kHz = 40kHz$. This is the fundamental reason for the adoption of sampling frequencies like $44.1kHz$ and $48kHz$, standards in Compact Disks (CD) and broadcast systems (radio and television), respectively.

Within this framework, musical notes can be characterized. The note still stands paradigmatic as the 'fundamental unit' of musical structures and, in practice, it can unfold into sounds that uphold other approaches\footnote{Music from the twentieth century enlarged this traditional comprehension of music. This occurred in the concert music, specially in the concrete, electronic and electroacoustic styles. On the last decade of the century, it became evident that popular music had also incorporated sounds without defined pitch and temporal organization, to name a few of simple characteristics.}. 
Notes are also convenient for another reason: the average listener -- and considerable part of the specialists -- presupposes rhythmic and pitch organization (made explicit in section~\ref{sec:notasMusic}) as fundamental musical properties, and these are developed in traditional musical theory in terms of notes.

@@

\nocite{*}
\bibliography{article}
\end{document}
